Synapse R API Tutorial
Analyzing and recording analysis done on external data from GEO(Gene Expression Omnibus)
========================================================


### 1. Load The Required Libraries
```{r}
suppressMessages(require(synapseClient))
suppressMessages(require(affy))
suppressMessages(require(simpleaffy))
suppressMessages(require(limma))
```


### Login to Synapse
```{r}
synapseLogin()
```


### 2. Create a Project on Synapse
```{r}
random_string <- paste(sample(LETTERS,4),collapse="")
proj_name <- paste('Demo_external_GEO_data_analysis',random_string,sep='-')
myProj <- Project(name=proj_name)
myProj <- synStore(myProj)
#onWeb(myProj)
folderId <- myProj$properties$id
```


Alternatively one could provide the FOLDER ID where the files could be stored
folderId <- ""


### 3. Reference a file stored in GEO and register with Synapse
```{r}
external_data_url <- "ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE4nnn/GSE4757/suppl/GSE4757_RAW.tar"
f <- File(path=external_data_url, 
          parentId=folderId, synapseStore=FALSE)
f <- synStore(f)

#f <- synGet(f$properties$id)
# f_location <- getFileLocation(f)

###  To avoid downloading data for demo I am using a local copy 
###  The users should do the actual download in order to get the data
### by uncommenting the above line f <- synGet(f$properties$id) & f_location <- getFileLocation(f)
f_location <- "/tmp/GSE4757_RAW.tar"
```


### 4. Open the zipped file and read its contents
```{r}
## Untar the local copy of the file
celDir <- file.path(tempdir(), "celfiles")
dir.create(celDir)
untar(f_location, exdir=celDir, tar="/usr/bin/tar")

## Read the microarray CEL files
ab <- ReadAffy(celfile.path=celDir)
scanDates <- as.Date(substr(ab@protocolData@data$ScanDate, 1, 8), format="%m/%d/%y")
ab <- ab[, order(scanDates)]
scanDates <- scanDates[order(scanDates)]
```


### 5. Lets look at the RAW pre-Normalized expression
```{r}
boxPath <- file.path(tempdir(), "gse4757-boxplot-prenormalization.png")
png(boxPath, width=600, height=400)
boxplot(log2(exprs(ab)), range=0, col=factor(scanDates), xlab="CEL File", ylab="log2(expr)", main="Pre-Normalization colored by Scan Date")
dev.off()
boxplot(log2(exprs(ab)), range=0, col=factor(scanDates), xlab="CEL File", ylab="log2(expr)", main="Pre-Normalization colored by Scan Date")
```


### 6. Push the plot to synapse
```{r}
boxFile <- synStore(File(path=boxPath, parentId=folderId), used=f)
#onWeb(boxFile)
```


### 7. Simple RMA Normalization
```{r}
eset <- rma(ab)
boxPathPost <- file.path(tempdir(), "gse4757-boxplot-postnormalization.png")
png(boxPathPost, width=600, height=400)
boxplot(exprs(eset), range=0, col=factor(scanDates), xlab="CEL File", ylab="log2(expr)", main="Post-Normalization (RMA) colored by Scan Date")
dev.off()
boxplot(exprs(eset), range=0, col=factor(scanDates), xlab="CEL File", ylab="log2(expr)", main="Post-Normalization (RMA) colored by Scan Date")
```

### 8. Push the normalized expression plot to Synapse
```{r}
boxFilePost <- synStore(File(path=boxPathPost, parentId=folderId), used=f)
#onWeb(boxFilePost)
```


### 9. Finally clean up the project (as this was a demo)
You can skip this step if you want to keep your project
```{r}
synDelete(myProj)
```

