# Getting started with Synapse using R

Import the synapse R client and log in to Synapse using locally stored credentials.
```{r}
library(synapseClient)
synapseLogin()
```


#### Create a Synapse Project to store your work:
```{r}
myProject = Project(name = "Demo Synapse Project")
myProject = synStore(myProject)
print(paste('Created a project with Synapse id', myProject$properties$id, sep = ' '))
```

#### Or, start with an existing project, such as the one created above:
```{r}
id = myProject$properties$id
rm(myProject)
myProject = synGet(id)
```

#### See your project online:
```{r}
onWeb(myProject)
```

#### Create a wiki for the project and put some text on it:
```{r}
placeholderText = "Place-holder text: Credibly innovate granular internal or organic sources whereas high standards in web-readiness. Energistically scale future-proof core competencies vis-a-vis impactful experiences. Dramatically synthesize integrated schemas with optimal networks."
wiki = WikiPage(owner=myProject, title="Analysis summary", markdown=placeholderText)
wiki = synStore(wiki)
```


#### Organize the project by creating folders within it:
```{r}
myFolderEntity = Folder(name = "plots", parentId = myProject$properties$id)
myFolderEntity = synStore(myFolderEntity)
```

#### See your project online:
```{r}
onWeb(myProject)
```

#### Create and upload a plot to the project:
```{r}
x = rnorm(500,mean=6,sd=4)
y = rnorm(500,mean=2,sd=3)
pdf(file="demo_file.pdf")
par(mfrow = c(1,2))
hist(x, col = "red", xlim = range(-10,15))
hist(y, col = "blue", xlim = range(-10,15))
dev.off()

plotFileEntity = File(path="demo_file.pdf", parentId=myFolderEntity$properties$id)
synSetAnnotations(plotFileEntity) = list(sampleType="iPSC", institution="FredHutch", protocol="A43.6")
plotFileEntity = synStore(plotFileEntity)
```


#### Add provenance describing how plot was created:
```{r}
plotFileEntity = synStore(plotFileEntity, executed='https://github.com/Sage-Bionetworks/synapseTutorials/blob/master/R/getting_started.Rmd')
```

#### View project on web:
```{r}
onWeb(plotFileEntity)
```
